---
title: Modifying Chromosome Nomenclature and Filtering for T~m~
output:
  html_notebook:
    theme: flatly
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
---

At this stage, the uniquely-mapped probes have been filtered for structure. They now have to be spaced evenly. This code attempts to do that, as well as evaluate the consistency of the spacing.

```{r }
library(tidyverse)
library(Gviz)
library(rafalib)
```

# Reading in the file

```{r message=FALSE}
bed_align <- read_tsv(file = "Tgfbr1_cleaned_probes_chr_sC.bed", col_names = FALSE)

colnames(bed_align) <- c("chromosome", "start", "end", "probe", "Tm")

bed_align
```

# Creating evenly-spaced probe sets

This code attempts to create multiple probe sets which are spaced at a distance of more than or equal to `min_probe_spacing`. Each probe set begins at a certain probe in the full probe collection (hereafter referred to as the "anchor probe"), and finds the next probe which is at least at a distance of `min_probe_spacing` from the first probe, continuing the process until the last viable probe fulfilling the criterion is reached. Picking different anchor probe results in a different probe set.

At the end, one probe set is to be picked, which can be used for region-specific extraction.

```{r }
x <- vector(mode = "list", length = (nrow(bed_align) - 1)) # initiating a blank list for holding all the probe sets

names(x) <- paste0(as.character(1:(nrow(bed_align) - 1)))

names(x) %>% head()
```


```{r }
min_probe_spacing <- 3000 # minimum space to be kept between end of first probe and start of next

for(i in 1:(nrow(bed_align) - 1)){
  
  j <- i
  
  a <- j+1
  
  while(a <= nrow(bed_align)){
    
    if((bed_align$start[a] - bed_align$end[j]) <= min_probe_spacing){
      a <- a+1
    } else {
      x[[as.character(i)]] <- append(x[[as.character(i)]], c("probe_1" = j, "probe_2" = a))
      j <- a
      a <- j+1
    }
    
  } 
}
```

```{r }
length(x)
```

Each element in the list contains indices referring to specific probes contained in the `bed_align` data frame. These indices are repeated twice; hence, they have to be deduplicated to create a list of unique indices.

```{r }
x$`1`
```

The probe set list also contains null elements, which have to be removed.

```{r }
x$`821`
```

```{r }
x <- purrr::compact(x)  # removing null list elements
```

Getting a unique set of probe indices.

```{r }
x <- sapply(x, unique)
x$`1`
```

Giving better names for the elements in the probe set list.

```{r }
names(x) <- paste0("probeset_", names(x))

names(x) %>% head()
```

Getting the number of probes present in each probe set.


```{r }
probeset_lengths <- sapply(x, length)

names(probeset_lengths) <- NULL

probeset_lengths %>% head()

probeset_lengths %>% table()
```

The probe sets with the most number of probes contain 20 probes each. Getting a modified collection of probe sets consisting of these probe sets.

```{r }
largest_probesets <- x[which(probeset_lengths == max(probeset_lengths))]

largest_probesets$probeset_1
```

# Getting probe spacing statistics

```{r }
probe_spacing <- vector(mode = "list", length = length(largest_probesets)) # creating null list for holding probe spacing values for all probe sets

names(probe_spacing) <- names(largest_probesets)

mean_probe_spacing <- vector(mode = "numeric", length = length(largest_probesets)) # creating null vector for holding probe spacing values for all probe sets

names(mean_probe_spacing) <- names(largest_probesets)

interquartile <- vector(mode = "numeric", length = length(largest_probesets))  # creating null vector for holding interquartile ranges of all probe sets

names(interquartile) <- names(largest_probesets)

for(i in seq_along(largest_probesets)){  # iterating along all the probe sets
  
  current_set_spacing <- vector(mode = "numeric", length = length(largest_probesets[[i]])) # create a null vector for holding distance between end of one probe and start of next probe (i.e., probe spacing) for current probe set
  
  for(j in 1:(length(current_set_spacing) - 1)){ # iterating along the probes in single probe set
    
    selected_probe_idx <- largest_probesets[[i]][j]  # index of selected probe in probe set
    
    next_probe_idx <- largest_probesets[[i]][j+1] # index of the next probe in probe set
    
    current_set_spacing[j] <- (bed_align$start[next_probe_idx] - bed_align$end[selected_probe_idx]) # start of next probe - end of selected probe
  }
  
  current_set_spacing <- current_set_spacing[-length(current_set_spacing)] # remove final element of current_set_spacing, which is a zero and is redundant
  
  probe_spacing[[i]] <- current_set_spacing # probe spacing for entire probe set
  
  mean_probe_spacing[i] <- mean(current_set_spacing)
  
  interquartile[i] <- IQR(current_set_spacing) # interquartile range of probe spacing of probe set
}
```


```{r }
probe_spacing$probeset_1
```

```{r }
mean_probe_spacing
```

```{r }
interquartile
```

```{r }
mypar(1, 1)

boxplot(mean_probe_spacing, main="Mean probe spacing for each set", horizontal=TRUE)
```

```{r }
boxplot(interquartile, main = "Interquartile ranges of probe spacing for each set", horizontal=TRUE)
```

```{r }
hist(interquartile, main = "Interquartile ranges of probe spacing for each set")
```

The first probeset seems satisfactory enough in terms of probe spacing criteria.

```{r }
spaced_probes <- bed_align[largest_probesets$probeset_1, ]
```

# Writing the selected spaced probe set to a file

```{r }
write_tsv(x = spaced_probes, path = "Tgfbr1_spaced_probes.bed", col_names = FALSE)
```

# Visualizing the spacing of the probe sets using Gviz

```{r }
spaced_probe_ranges <- GRanges(seqnames = spaced_probes$chromosome, ranges = IRanges(start = spaced_probes$start, end = spaced_probes$end), probe = spaced_probes$probe, Tm = spaced_probes$Tm)

probe_track <- Gviz::AnnotationTrack(range = spaced_probe_ranges, chromosome = 4, name  = "probes")

gtrack <- GenomeAxisTrack()

tgfbr1_bounds <- c("start" = 47353222, "end" = 47414926)

tgfbr1_range <- GRanges(seqnames = c("chr4"), ranges = IRanges (start = tgfbr1_bounds["start"], end = tgfbr1_bounds["end"]))

tgfbr1_track <- AnnotationTrack(range = tgfbr1_range, name = "Tgfbr1")
```

```{r }
plotTracks(trackList =  list(probe_track, gtrack, tgfbr1_track))
```

```{r }
spaced_probes$Tm %>% summary()
```


